detect patches

### 2024-01-31 scripts/classify_patches.Rmd
adding skeletons for grabbing occurrence and abundance data

lapply(f[4:20], FUN = landscapR, pout = 'patch_metrics', pRout = 'fine_patches')

@@

# nearest neighbor distance (enn_mn) - (Class) {lsm_c_enn_mn} # note other method of aggregation can be used too!
## (modification) log(enn_mn), sqrt(enn_mn)
### Expectation: the number of colonized predicted patches increases with increasing distance.
#### why: this species is adept at long distance dispersal
##### test: glm(colonization(0|1) ~ enn_mn,  family = binomial)

@@

>>>>

lapply(f[21:335], FUN = landscapR, pout = 'patch_metrics', pRout = 'fine_patches')

@@

# nearest neighbor distance (enn_mn) - (Class) {lsm_c_enn_mn} # note other method of aggregation can be used too!
## (modification) log(enn_mn), sqrt(enn_mn)
### Expectation: the proportion of predicted patches which are colonized decreases with increaing mean distance between patches
#### why: further distances between suitable habitats prevent colonization and recolonization of habitat.
##### test: glm(prcnt_patches_colonized ~ enn_mn,  family = (quasi)binomial)

@@

To determine whether a patch is occupied the original training data are matched as an attribute to the fine patches.

f <- file.path('../results/fine_patches', list.files('../results/fine_patches'))

recs <- st_read('../data/raw/occurrence/combined_records') %>% 
  filter(PtType == 1)

patchTaggR <- function(x){

  r <- terra::rast(f)
  r_taxon <- gsub('[.]tif', '', basename(f))
  occs <- dplyr::filter(recs, taxon == r_taxon) |>
    dplyr::mutate(ID = 1:nrow(.)) |>
    terra::vect()

  occ_pts <- terra::extract(r, occs, bind = TRUE) |>
    sf::st_as_sf()

  # if pts were NA, buffer to 90m, many occupied areas were 'cut' away by the 
  # hydrologic mapping. 
  occ_poly <- dplyr::filter(occ_pts, is.na(RASTER)) |>
    sf::st_buffer(occ_pts, 120) # we cut out 90, but let's go a bit further for corner cases.
  occ_poly <- terra::extract(r, occ_poly, bind = TRUE, weights = TRUE) |>
    sf::st_as_sf() |>
    sf::st_centroid()  |>
    dplyr::group_by(ID) |> # now make sure we get the patch with the 'most' overlap with the polygon
    dplyr::slice_max(WEIGHT) |>
    dplyr::drop_na(RASTER) # these values actually NA.

  occ_pops <- dplyr::bind_rows( # a dataframe of all raster patch ID's with populations known from within
    # the last few decades in them. 
    dplyr::drop_na(occ_pts), occ_poly) |>
    dplyr::distinct(RASTER, .keep_all = TRUE) # this is multiple pts per patch - not neccessary for us.

  write.csv(occ_pops, row.names = F,
            file.path('../results/occupiedPatches', paste0(r_taxon, '.csv')))

p1 <- '../data/raw/2021_AIM_Terrestrial/AIMTerrestrial9-1-22.gdb'
st_layers(dsn = p1)

AIM_points <- st_read(dsn = p1, 
        layer = 'TerrADat', quiet = T) %>%
  select(PrimaryKey, DateVisited, geometry = Shape) %>% 
  st_transform(5070)

lpi <- spp_richness <- st_read(dsn = p1, 
        layer = 'tbl LINE POINT INTERCEPT', quiet = T) %>% 
  select(SpeciesList, PrimaryKey) %>% 
  right_join(AIM_points, by = 'PrimaryKey') %>% 
  st_as_sf()  %>%

  # NOW CALCULATE THE PERCENT COVER OF EACH SPECIES FROM LPI #
  # ...

  mutate(DateVisited = as.Date(DateVisited), .before = geometry) %>% 
  arrange(SpeciesList) %>% 
  right_join(., select(targets, taxon, symbol) %>% 
               distinct(), by = c('SpeciesList' = 'symbol' ))

# add the species names from USDA PLANTS CODES

# ...

AIM_abundance_retrievR <- function(x, aim_tables){
